A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-19 06-appendix.tex
Adiciona ao apêndice prova (incompl.) da terminação NST

% \makeatletter
% \newcommand\case{\@startsection{paragraph}{4}{\z@}%
%    {-.5\baselineskip \@plus -2\p@ \@minus -.2\p@}%
%    {-3.5\p@}%
%    {\@parfont}}
%  \def\@parfont{\itshape}
% \makeatother

@@

\begin{theorem}\label{thm.det-out-pop-n}
  If~$\delta\out[n]\delta_1$ and~$\delta\out[n]\delta_2$
  then~$\delta_1=\delta_2$.
\end{theorem}
\begin{proof}
  By induction on~$n$.
  %%
  The theorem is trivially true for~$n=0$ and follows directly from
  Lemma~\ref{lem.det-out} for~$n=1$.  (Note that for~$n>1$ the theorem is
  vacuously true for~$\outpush$ transitions---by the format of
  rules~\R{push} and~\R{pop}, transitions~$\outpush$ cannot be applied more
  than once in a row and cannot occur after a~$\outpop$ transition.)

  Suppose
  \[
    \delta\out[1]\delta_1'\out[n-1]\delta_1
    \quad\text{and}\quad
    \delta\out[1]\delta_2'\out[n-1]\delta_2,
  \]
  for some~$n>1$ and~$\delta_1'$, $\delta_2'\in\Delta$.
  %%
  By Lemma~\ref{lem.det-out}, $\delta_1'=\delta_2'$.  By induction
  hypothesis, $\delta_1=\delta_2$.\qedhere
\end{proof}

@@

  \begin{case}
    $p=\ceu{\Ifelse{\Mem(\Id)}{p'}{p''}}$.  There are two subcases.
    \begin{subcase}
      $\eval(\ceu{\Mem(\Id)})$ is true.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{if-true}.
      Hence~$p_1=p_2=p'$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $\eval(\ceu{\Mem(\Id)})$ is false.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{if-false}.
      Hence~$p_1=p_2=p''$ and~$e_1=e_2=\nil$.
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p';\,p''}$.  There are three subcases.
    \begin{subcase}
      $p'=\ceu{\Nop}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-nop}.
      Hence~$p_1=p_2=p''$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'=\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-brk}.
      Hence~$p_1=p_2=\ceu{\Break}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-adv}.
      Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
      \begin{align*}
        \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
        \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
      \end{align*}
      for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By induction
      hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
      Hence~$p_1=\ceu{p_1';p''}=\ceu{p_2';p''}=p_2$ and~$e_1=e_1'=e_2'=e_2$.
    \end{subcase}
  \end{case}

@@

  \begin{case}
    $p=\ceu{p'\AtLoop{p''}}$.  There are three subcases.
    \begin{subcase}
      $p'=\ceu{\Nop}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-nop}.
      Hence~$p_1=p_2=\ceu{\Loop{p''}}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'=\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-break}.
      Hence~$p_1=p_2=\ceu{\Nop}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-adv}.
      Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
      \begin{align*}
        \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
        \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
      \end{align*}
      for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By induction
      hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
      Hence~$p_1=\ceu{p_1'\AtLoop{p''}}=\ceu{p_2'\AtLoop{p''}}=p_2$
      and~$e_1=e_1'=e_2'=e_2$.
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p'\And{p''}}$.
    %%
    Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-expd}.
    Hence~$p_1=p_2=\ceu{{p'}\And{(\CanRun(n);\,p'')}}$ and~$e_1=e_2=\nil$.
  \end{case}

  \begin{case}
    $p=\ceu{p'\AtAnd{p''}}$.  There are two subcases.
    \begin{subcase}
      $\isblocked(p',n)$~is false.  There are three subcases.
      \begin{subsubcase}
        $p'=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-nop1}.
        Hence~$p_1=p_2=p''$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-brk1}
        $p'=\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-brk1}.
        Hence~$p_1=p_2=\ceu{\clear(p'');\Break}$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-adv1}
        $p'\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-adv1}.
        Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
        \begin{align*}
          \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
          \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
        \end{align*}
        for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By induction
        hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
        Hence~$p_1=\ceu{{p_1'}\And{p''}}=\ceu{{p_2'}\And{p''}}=p_2$
        and~$e_1=e_1'=e_2'=e_2$.
      \end{subsubcase}
    \end{subcase}
    \begin{subcase}
      $\isblocked(p',n)$~is true.  There are three subcases.
      \begin{subsubcase}
        $p''=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-nop2}.
        Hence~$p_1=p_2=p'$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-brk2}
        $p''=\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-brk2}.
        Hence~$p_1=p_2=\ceu{\clear(p');\Break}$ and~$e_1=e_2=\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-adv2}
        $p''\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-adv2}.  Thus there
        are derivations~$\pi_1''$ and~$\pi_2''$ such that
        \begin{align*}
          \pi_1''&\Vdash\<p'',n,\nil>\nst\<p_1'',n,e_1''>\\
          \pi_2''&\Vdash\<p'',n,\nil>\nst\<p_2'',n,e_2''>
        \end{align*}
        for some~$p_1'',p_2''\in\P$ and~$e_1'',e_2''\in\E$.  By induction
        hypothesis, $p_1''=p_2''$ and~$e_1''=e_2''$.
        Hence~$p_1=\ceu{{p'}\And{p_1''}}=\ceu{{p'}\And{p_2''}}=p_2$
        and~$e_1=e_1''=e_2''=e_2$.
      \end{subsubcase}
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p'\Or{p''}}$.
    %%
    Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-expd}.
    Hence~$p_1=p_2=\ceu{{p'}\Or{(\CanRun(n);\,p'')}}$ and~$e_1=e_2=\nil$.
  \end{case}

  \begin{case}
    $p=\ceu{p'\AtOr{p''}}$.  There are two subcases.
    \begin{subcase}
      $\isblocked(p',n)$~is false.  There are three subcases.
      \begin{subsubcase}
        $p'=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-nop1}.
        Hence~$p_1=p_2=\clear(p'')$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}
        $p'=\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-brk1}.
      \end{subsubcase}
      \begin{subsubcase}
        $p'\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-adv1}.
      \end{subsubcase}
    \end{subcase}
    \begin{subcase}
      $\isblocked(p',n)$~is true.  There are three subcases.
      \begin{subsubcase}
        $p''=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-nop1}.
        Hence~$p_1=p_2=\clear(p')$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}
        $p''=\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-brk2}.
      \end{subsubcase}
      \begin{subsubcase}
        $p''\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-adv2}.
        %%
        \qedhere
      \end{subsubcase}
    \end{subcase}
  \end{case}
\end{proof}

@@

\begin{definition}\label{def.Hnst}
  A description~$\delta=\<p,n,e>$ is \emph{nested-irre\-ducible}
  iff~$e\ne\nil$ or~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p,n)$~is
  true.  Nested-irreducible descriptions serve as normal forms for~$\nst$
  transitions: they embody the result of an exhaustive number of~$\nst$
  transitions.  We will write~$\delta_\Hnst$ to indicate that
  description~$\delta$ is nested-irreducible.
\end{definition}

@@

\begin{lemma}\label{lem.props-nst-n}
  If~$\<p_1,n,e>\nst[n]\<p_1',n,e'>$ then, for any~$p_2$:
  \begin{enumerate}[(a)]
  \item\label{lem.props-nst-n.a}
    $\<\ceu{p_1;\,p_2},n,e>\nst[n]\<p_1';p_2,n,e'>$;
    %%
  \item\label{lem.props-nst-n.b}
    $\<\ceu{p_1\AtLoop{p_2}},n,e>\nst[n]\<\ceu{p_1'\AtLoop{p_2}},n,e'>$;
    %%
  \item\label{lem.props-net-n.c}
    $\<\ceu{{p_1}\AtAnd{p_2}},n,e>\nst[n]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>$;
    %%
  \item\label{lem.props-net-n.d}
    $\<\ceu{{p_1}\AtOr{p_2}},n,e>\nst[n]\<\ceu{{p_1}'\AtOr{p_2}},n,e'>$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on~$n$.
  %%
  \begin{enumerate}[(a)]
  \item The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    directly from~\R{seq-adv} for~$n=1$.  Suppose
    \begin{equation}
      \label{lem.props-nst-n.a.eq1}
      \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>,
    \end{equation}
    for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible, i.e.,
    $e=\nil$ and~$p\ne{\ceu{\Nop},\ceu{\Break}}$ and~$\isblocked(p_1,n)$ is
    false.  By~\eqref{lem.props-nst-n.a.eq1} and by~\R{seq-adv},
    \begin{equation}
      \label{lem.props-nst-n.a.eq2}
      \<\ceu{p_1;\,p_2},n,e>\nst[1]\<\ceu{p_1'';\,p_2},n,e''>.
    \end{equation}
    From~\eqref{lem.props-nst-n.a.eq1}, by induction hypothesis,
    \begin{equation}
      \label{lem.props-nst-n.a.eq3}
      \<\ceu{p_1'';\,p_2},n,e''>\nst[n-1]\<\ceu{p_1';\,p_2},n,e'>.
    \end{equation}
    From~\eqref{lem.props-nst-n.a.eq2} and~\eqref{lem.props-nst-n.a.eq3},
    \[
      \<\ceu{p_1;\,p_2},n,e>\nst[n]\<\ceu{p_1';\,p_2},n,e'>.
    \]
  \end{enumerate}
\end{proof}

\begin{theorem}\label{thm.term-nst-*}
  For any~$\delta$ there is a~$\delta'_\Hnst$ such
  that~$\delta\nst[*]\delta'_\Hnst$.
\end{theorem}
\begin{proof}
  By induction on the structure of programs.
  %%
  Let~$\delta=\<p,n,\nil>$.  The theorem is trivially true if~$\delta$ is
  nested-irreducible itself, as by definition~$\delta\nst[0]\delta_\Hnst$.
  Suppose that is not the case.  Then, depending on the structure of~$p$,
  there are~11 possibilities.  In each one of them, we show that
  such~$\delta'_\Hnst$ indeed exists.
  \begin{case}
    $p=\ceu{\Mem(\Id)}$.
    %%
    Then, by~\R{mem},
    \[
      \<\ceu{\Mem(\Id)},n,\nil>\nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
    \]
  \end{case}

>>>>

\subsection*{Determinism}

@@

\begin{theorem}\label{thm.det-out-pop-n}
  If~$\delta\out[n]\delta_1$ and~$\delta\out[n]\delta_2$
  then~$\delta_1=\delta_2$.
\end{theorem}
\begin{proof}
  By induction on~$n$.
  %%
  The theorem is trivially true for~$n=0$ and follows directly from
  Lemma~\ref{lem.det-out} for~$n=1$.  (Note that for~$n>1$ the theorem is
  vacuously true for~$\outpush$ transitions---by the format of
  rules~\R{push} and~\R{pop}, transitions~$\outpush$ cannot be applied more
  than once in a row and cannot occur after a~$\outpop$ transition.)
  %%
  Suppose~$\delta\out[1]\delta_1'\out[n-1]\delta_1$
  and~$\delta\out[1]\delta_2'\out[n-1]\delta_2$, for some~$n>1$
  and~$\delta_1'$, $\delta_2'\in\Delta$.
  %%
  By Lemma~\ref{lem.det-out}, $\delta_1'=\delta_2'$.  By the induction
  hypothesis, $\delta_1=\delta_2$.\qedhere
\end{proof}

@@

  \begin{case}
    $p=\ceu{\Ifelse{\Mem(\Id)}{p'}{p''}}$.  There are two subcases.
    \begin{subcase}
      $\eval(\ceu{\Mem(\Id)})$~is true.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{if-true}.
      Hence~$p_1=p_2=p'$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $\eval(\ceu{\Mem(\Id)})$ is false.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{if-false}.
      Hence~$p_1=p_2=p''$ and~$e_1=e_2=\nil$.
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p';\,p''}$.  There are three subcases.
    \begin{subcase}
      $p'=\ceu{\Nop}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-nop}.
      Hence~$p_1=p_2=p''$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'=\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-brk}.
      Hence~$p_1=p_2=\ceu{\Break}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{seq-adv}.
      Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
      \begin{align*}
        \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
        \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
      \end{align*}
      for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By the induction
      hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
      Hence~$p_1=\ceu{p_1';p''}=\ceu{p_2';p''}=p_2$ and~$e_1=e_1'=e_2'=e_2$.
    \end{subcase}
  \end{case}

@@

  \begin{case}
    $p=\ceu{p'\AtLoop{p''}}$.  There are three subcases.
    \begin{subcase}
      $p'=\ceu{\Nop}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-nop}.
      Hence~$p_1=p_2=\ceu{\Loop{p''}}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'=\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-break}.
      Hence~$p_1=p_2=\ceu{\Nop}$ and~$e_1=e_2=\nil$.
    \end{subcase}
    \begin{subcase}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then~$\pi_1$ and~$\pi_2$ are instances of~\R{loop-adv}.
      Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
      \begin{align*}
        \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
        \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
      \end{align*}
      for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By the induction
      hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
      Hence~$p_1=\ceu{p_1'\AtLoop{p''}}=\ceu{p_2'\AtLoop{p''}}=p_2$
      and~$e_1=e_1'=e_2'=e_2$.
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p'\And{p''}}$.
    %%
    Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-expd}.
    Hence~$p_1=p_2=\ceu{{p'}\AtAnd{(\CanRun(n);\,p'')}}$ and~$e_1=e_2=\nil$.
  \end{case}

  \begin{case}
    $p=\ceu{p'\AtAnd{p''}}$.  There are two subcases.
    \begin{subcase}
      $\isblocked(p',n)$~is false.  There are three subcases.
      \begin{subsubcase}
        $p'=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-nop1}.
        Hence~$p_1=p_2=p''$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-brk1}
        $p'=\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-brk1}.
        Hence~$p_1=p_2=\ceu{\clear(p'');\Break}$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-adv1}
        $p'\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-adv1}.
        Thus there are derivations~$\pi_1'$ and~$\pi_2'$ such that
        \begin{align*}
          \pi_1'&\Vdash\<p',n,\nil>\nst\<p_1',n,e_1'>\\
          \pi_2'&\Vdash\<p',n,\nil>\nst\<p_2',n,e_2'>
        \end{align*}
        for some~$p_1',p_2'\in\P$ and~$e_1',e_2'\in\E$.  By the induction
        hypothesis, $p_1'=p_2'$ and~$e_1'=e_2'$.
        Hence~$p_1=\ceu{{p_1'}\And{p''}}=\ceu{{p_2'}\And{p''}}=p_2$
        and~$e_1=e_1'=e_2'=e_2$.
      \end{subsubcase}
    \end{subcase}
    \begin{subcase}
      $\isblocked(p',n)$~is true.  There are three subcases.
      \begin{subsubcase}
        $p''=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-nop2}.
        Hence~$p_1=p_2=p'$ and~$e_1=e_2\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-brk2}
        $p''=\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-brk2}.
        Hence~$p_1=p_2=\ceu{\clear(p');\Break}$ and~$e_1=e_2=\nil$.
      \end{subsubcase}
      \begin{subsubcase}\label{lem.det-nst.and-adv2}
        $p''\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{and-adv2}.  Thus there
        are derivations~$\pi_1''$ and~$\pi_2''$ such that
        \begin{align*}
          \pi_1''&\Vdash\<p'',n,\nil>\nst\<p_1'',n,e_1''>\\
          \pi_2''&\Vdash\<p'',n,\nil>\nst\<p_2'',n,e_2''>
        \end{align*}
        for some~$p_1'',p_2''\in\P$ and~$e_1'',e_2''\in\E$.  By the
        induction hypothesis, $p_1''=p_2''$ and~$e_1''=e_2''$.
        Hence~$p_1=\ceu{{p'}\And{p_1''}}=\ceu{{p'}\And{p_2''}}=p_2$
        and~$e_1=e_1''=e_2''=e_2$.
      \end{subsubcase}
    \end{subcase}
  \end{case}

  \begin{case}
    $p=\ceu{p'\Or{p''}}$.
    %%
    Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-expd}.
    Hence~$p_1=p_2=\ceu{{p'}\AtOr{(\CanRun(n);\,p'')}}$ and~$e_1=e_2=\nil$.
  \end{case}

  \begin{case}
    $p=\ceu{p'\AtOr{p''}}$.  There are two subcases.
    \begin{subcase}
      $\isblocked(p',n)$~is false.  There are three subcases.
      \begin{subsubcase}
        $p'=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-nop1}.
        Hence~$p_1=p_2=\clear(p'')$ and~$e_1=e_2=\nil$.
      \end{subsubcase}
      \begin{subsubcase}
        $p'=\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-brk1}.
      \end{subsubcase}
      \begin{subsubcase}
        $p'\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-adv1}.
      \end{subsubcase}
    \end{subcase}
    \begin{subcase}
      $\isblocked(p',n)$~is true.  There are three subcases.
      \begin{subsubcase}
        $p''=\ceu{\Nop}$.
        %%
        Then~$\pi_1$ and~$\pi_2$ are instances of~\R{or-nop1}.
        Hence~$p_1=p_2=\clear(p')$ and~$e_1=e_2=\nil$.
      \end{subsubcase}
      \begin{subsubcase}
        $p''=\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-brk2}.
      \end{subsubcase}
      \begin{subsubcase}
        $p''\ne\ceu{\Nop},\ceu{\Break}$.
        %%
        Similar to Case~\ref{lem.det-nst.and-adv2}.
        %%
        \qedhere
      \end{subsubcase}
    \end{subcase}
  \end{case}
\end{proof}

@@

\subsection*{Termination}

\begin{definition}\label{def.Hnst}
  A description~$\delta=\<p,n,e>$ is \emph{nested-irre\-ducible}
  iff~$e\ne\nil$ or~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p,n)$~is
  true.  Nested-irreducible descriptions serve as normal forms for~$\nst$
  transitions: they embody the result of an exhaustive number of~$\nst$
  applications.  We will write~$\delta_\Hnst$ to indicate that
  description~$\delta$ is nested-irreducible.
\end{definition}

@@

\begin{lemma}\label{lem.props-nst-n}
  If~$\<p_1,n,e>\nst[n]\<p_1',n,e'>$ then, for any~$p_2$:
  \begin{enumerate}[(a)]
  \item\label{lem.props-nst-n.a}
    $\<\ceu{p_1;\,p_2},n,e>\nst[n]\<p_1';p_2,n,e'>$;
    %%
  \item\label{lem.props-nst-n.b}
    $\<\ceu{p_1\AtLoop{p_2}},n,e>\nst[n]\<\ceu{p_1'\AtLoop{p_2}},n,e'>$;
    %%
  \item\label{lem.props-nst-n.c}
    $\<\ceu{{p_1}\AtAnd{p_2}},n,e>\nst[n]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>$;
    %%
  \item\label{lem.props-nst-n.d}
    $\<\ceu{{p_1}\AtOr{p_2}},n,e>\nst[n]\<\ceu{{p_1}'\AtOr{p_2}},n,e'>$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on~$n$.
  %%
  \begin{enumerate}[(a)]
  \item The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    directly from~\R{seq-adv} for~$n=1$.  Suppose
    \begin{equation}
      \label{lem.props-nst-n.a.eq1}
      \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    \end{equation}
    for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible, i.e.,
    $e=\nil$ and~$p\ne{\ceu{\Nop},\ceu{\Break}}$ and~$\isblocked(p_1'',n)$
    is false.  By~\eqref{lem.props-nst-n.a.eq1} and by~\R{seq-adv},
    \begin{equation}
      \label{lem.props-nst-n.a.eq2}
      \<\ceu{p_1;\,p_2},n,e>\nst[1]\<\ceu{p_1'';\,p_2},n,e''>\,.
    \end{equation}
    From~\eqref{lem.props-nst-n.a.eq1}, by the induction hypothesis,
    \begin{equation}
      \label{lem.props-nst-n.a.eq3}
      \<\ceu{p_1'';\,p_2},n,e''>\nst[n-1]\<\ceu{p_1';\,p_2},n,e'>\,.
    \end{equation}
    From~\eqref{lem.props-nst-n.a.eq2} and~\eqref{lem.props-nst-n.a.eq3},
    \[
      \<\ceu{p_1;\,p_2},n,e>\nst[n]\<\ceu{p_1';\,p_2},n,e'>\,.
    \]

  \item Similar to Case~\eqref{lem.props-nst-n.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{loop-adv} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-n.b.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-n.b.eq1} and by~\R{loop-adv},
    % \begin{equation}
    %   \label{lem.props-nst-n.b.eq2}
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[1]\<\ceu{p_1''\Loop{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.b.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-n.b.eq3}
    %   \<\ceu{p_1''\AtLoop{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.b.eq2} and~\eqref{lem.props-nst-n.b.eq3},
    % \[
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[n]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to Case~\eqref{lem.props-nst-n.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{and-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-n.c.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-n.c.eq1} and by~\R{and-adv1},
    % \begin{equation}
    %   \label{lem.props-nst-n.c.eq2}
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtAnd{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.c.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-n.a.eq3}
    %   \<\ceu{{p_1''}\AtAnd{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.c.eq2} and~\eqref{lem.props-nst-n.c.eq3},
    % \[
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to Case~\eqref{lem.props-nst-n.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{or-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-n.d.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-n.d.eq1} and by~\R{or-adv1},
    % \begin{equation}
    %   \label{lem.props-nst-n.d.eq2}
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtOr{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.d.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-n.a.eq3}
    %   \<\ceu{{p_1''}\AtOr{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-n.d.eq2} and~\eqref{lem.props-nst-n.d.eq3},
    % \[
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \]
    %%
  \end{enumerate}
\end{proof}

The syntactic restrictions discussed in Section~\ref{?}, regarding the body
of~$\ceu{\Fin}$ and~$\ceu{\Loop}$ expressions, is formalized by the
following assumptions.

\begin{assumption}\label{ass.term-nst-fin}
  If~$\delta=\<\ceu{\clear(p)},n,\nil>$ then there is a
  description~$\delta'=\<\ceu{\Nop},n,\nil>$ such
  that~$\delta\nst[*]\delta'$.
\end{assumption}

\begin{assumption}\label{ass.term-nst-loop}
  If~$\delta=\<\ceu{\Loop{p}},n,\nil>$ then there is a
  description~$\delta'=\<p',n,e>$ such that~$\delta\nst[*]\delta'$ where
  either~$p'=\ceu{{\Break}\AtLoop{p}}$ or~$\isblockedext(p',n)$.
\end{assumption}
